Multi-fluids for Representing Sub-filter-scale Convection

### 2020-04-11 submit2/makeFigs/RCE_singleColumn.tex
First complete draft with Hilary's revisions and responses to reviewers.

%% LyX 2.2.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,times]{extarticle}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\usepackage{array}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}

\usepackage{todonotes}
%\usepackage{afterpage}

%\usepackage[switch]{lineno}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\usepackage{moreverb}

\usepackage{flushend}

@@

\begin{document}

\title{Multi-fluids for Representing Sub-filter-scale Buoyant Convection}

\author{Hilary Weller, William McIntyre, Daniel Shipley and Peter Clark}
\maketitle
\begin{abstract}
Multi-fluid modelling is a promising new method of representing sub
and near grid-scale convection allowing for consistent and accurate
treatment of net mass transport by convection, non-equilibrium and
turbulence. The air is partitioned into convecting and non-convecting
fluids each with their own velocity, temperature and constituents
so that separate equations of motion can be solved for each fluid,
thus representing features that cannot be resolved in a single fluid
at coarse resolution.

This paper presents multi-fluid Boussinesq equations for representing
under resolved dry convection. We present a new model for entrainment
and detrainment based on divergence which leads to excellent representation
of the convective area fraction. Previous multi-fluid modelling of
convection has assumed that all fluids share the same pressure. We
show that this is a bad assumption and we propose and validate a model
for the pressure difference between the fluids based on divergence.

Two vertical slice test cases are used to validate the multi-fluid
model; a buoyant rising bubble and radiative convective equilibrium
are simulated at high resolution and conditionally averaged based
on the sign of the vertical velocity. The multi-fluid model reproduces
the mean properties of the rising and falling fluids in a single column. 
\end{abstract}

\section{Introduction}

Slow progress has been made improving the parameterisation of sub-grid-scale
convection since \citet{AS74,Tied89,GR90,KF90}. Convection is still
one of the weakest aspects of large scale models of the atmosphere
\citep{HPB+14,SAB+13,ipcc41}. Improvements have been made relaxing
the quasi-equilibrium assumption \citep{PR98,GG05,Par14}, including
stochasticity \citep{PC08} and allowing net mass transport by convection
\citep{KB08,MB19}. These developments have enabled limited simulation
in the grey zone, where convection is partially resolved. However
there are still systematic biases implying that improvements should
still be possible.

\citet{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport correctly and consistently
with the same numerics and physics controlling thermals, the stable
environment and downdraught dynamics. \citet{Yano14} derived similar
equations for multiple segments and showed how this is a generalisation
of the mass-flux formulation. The \citet{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme also uses similar
equations but \citet{TKP+18} didn't propose a suitable numerical
method for solving the equations. Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citet{TEB19} and \citet{WM19} presented solutions
of the multi-fluid equations assuming two fluids that share the same
pressure (or that pressure differences are predominantly due to drag
between the fluids). \citet{TEB19} showed that these equations (without
drag) are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citet{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable and in this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}).

This paper presents the multi-fluid Boussinesq equations for representing
sub-grid-scale convection including a parameterisation of the pressure
difference between fluids (section \ref{subsec:fluidPressure}) and
a new model of entrainment and detrainment (section \ref{subsec:Sij}).
The numerical method for solving the equations is described in section
\ref{sec:numerics}. We compare results of resolved dry convection
in a single-fluid two dimensional Boussinesq model with a two-fluid,
single column model in section \ref{sec:results} using two vertical
slice test cases.

\section{The Multi-fluid Boussinesq Equations}

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \citep[eg those described in][]{WM19} carry
over to the Boussinesq equations; convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore for simplicity we focus on the
multi-fluid Boussinesq equations without background stratification:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\nabla\cdot\sigma_{i}\mathbf{u}_{i} & = & \sum_{j\ne i}\left\{ \sigma_{j}S_{ji}-\sigma_{i}S_{ij}\right\} \label{eq:sigma}\\
\frac{D_{i}\mathbf{u}_{i}}{Dt}+\nabla P_{i} & = & b_{i}\mathbf{k}+\nu\nabla^{2}\mathbf{u}_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(\mathbf{u}_{ji}^{T}-\mathbf{u}_{i}\right)-S_{ij}(\mathbf{u}_{ij}^{T}-\mathbf{u}_{i})-\mathbf{D}_{ij}\right\} \label{eq:mom}\\
\frac{D_{i}b_{i}}{Dt} & = & \alpha\nabla^{2}b_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(b_{ji}^{T}-b_{i}\right)-S_{ij}\left(b_{ij}^{T}-b_{i}\right)\right\} \label{eq:b}\\
\sum_{i}\nabla\cdot\sigma_{i}\mathbf{u}_{i} & = & \nabla\cdot\sum_{i}\sigma_{i}\mathbf{u}_{i}=0\label{eq:divFree}\\
\sum_{i}\sigma_{i} & = & 1\label{eq:sumOne}
\end{eqnarray}
with variables defined in table \ref{tab:defns}.

\begin{table}[h]
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)\tabularnewline
$P_{i}$ & Non-hydrostatic pressure of fluid $i$ $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$)\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$\tabularnewline
$b$ & Buoyancy defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$\gamma$ & Coefficient for setting the pressure local for each fluid ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
\end{tabular}

\caption{Definitions of variables for the multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

\subsection{Pressure of each fluid \label{subsec:fluidPressure}}

\citet{TEB19} and \citet{WM19} assumed that each co-located fluid
shared the same pressure, leading to unstable equations that \citet{TEB19}
stabilised using diffusion of vertical velocity and \citet{WM19}
stabilised using mass exchanges or drag between the fluids. \citet{TEB19}
achieved realistic results in single column experiments of convective
boundary layer but the amount of diffusion needed would spoil simulations
in the free troposphere. The stabilisation used by \citet{WM19} meant
that the two fluids tended to move as one which defeats the purpose
of using multi-fluid modelling.

The instabilities reported by \citet{WM19} showed growing divergence
in a fluid with vanishing volume fraction. It is mass convergence
rather than velocity convergence that leads to pressure anomalies
that remove the convergence. Therefore if the volume (or mass) fraction
of fluid is vanishingly low then convergence cannot lead to an anomaly
in the total pressure.

We will assume that pressure in each fluid is given by a parameterised
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}

\citet{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating fluidised beds of granular material
\citep[eg.][]{EHM92} and two compressible fluids \citep[eg.][]{SA99}.
In these methods, advection equations are written for the volume fraction,
$\sigma_{i}$, and conservation equations are written for mass fraction,
$\sigma_{i}\rho_{i}$. In our notation, ignoring transfers between
fluids, the \citet{BN86} equations would be:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)\label{eq:BNcompaction}\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}-\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

\citet{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citet{BN86} and
\citet{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot\mathbf{u}_{i}\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a free coefficient. Dimensional analysis in section
\ref{subsec:dimAnal} will inform the choice of $\gamma$ and sensitivity
to $\gamma$ will be evaluated in section \ref{sec:results}.

\subsection{Entrainment and Detrainment \label{subsec:Sij}}

We consider dry air without turbulence so use the dynamic entrainment
described by \citet{HC51,AK67,DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise}
\end{cases}
\end{equation}
which is necessary to stop air in one fluid from accumulating when
vertical motion ceases (at the top of a rising plume or at the bottom
of descending air). We will also test a common form of lateral entrainment
\citep{DBF+13}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}\implies S_{01}=w_{1}\frac{0.2}{r_{c}}
\end{equation}
where $\varepsilon$ is the well know fractional entrainment rate
and $r_{c}$ is the cloud or plume radius.

\subsection{Drag in the Momentum Equation\label{subsec:drag}}

Pressure differences between the fluids can lead to form drag which
is parameterised following \citet{SW69,RC15,WM19} as: 
\begin{equation}
\mathbf{D}_{ij}=\frac{\sigma_{j}C_{D}}{r_{c}}|\mathbf{u}_{i}-\mathbf{u}_{j}|\left(\mathbf{u}_{i}-\mathbf{u}_{j}\right)\label{eq:dragBubble}
\end{equation}
where $C_{D}$ is a drag coefficient, $r_{c}$ is a cloud radius.
Sensitivity to $C_{D}/r_{c}$is explored in section \ref{sec:results}.

\subsection{The Buoyancy and the Momentum of the Mass that is Transferred \label{subsec:transferProperties}}

The different fluids may not be well mixed so the fluid transferred
may not have the mean properties of the fluid it is leaving \citep[as was assumed by][]{WM19}.
In fact, the most buoyant air should be transferred from fluid 0 to
fluid 1 and vice versa, and the air with least downward momentum should
be transferred from 0 to 1 and vice versa. The properties of the fluids
transferred should depend on the modelling of sub-grid-scale variability
which is beyond the scope of this paper. In section \ref{sec:results}
we conditionally average high resolution solutions based on the sign
of $w$ so that $w<0$ air is in fluid 0 and $w>0$ air in fluid 1.
Therefore the air that is transferred at the interface will have $w=0$.
For two dimensional multi-fluid simulations we assume that the horizontal
velocity transferred is equal to the mean horizontal velocity of the
fluid transferred from. Therefore
\begin{equation}
\mathbf{u}_{ij}^{T}=\begin{pmatrix}u_{i}\\
v_{i}\\
0
\end{pmatrix}.
\end{equation}
For buoyancy, following \citet{TEB19}, we use:
\begin{equation}
b_{ij}^{T}=\theta_{b}b_{i}+(1-\theta_{b})b_{j}
\end{equation}
and present results using $\theta_{b}=\frac{1}{2}$ and $\theta_{b}=1$.
Note that $\theta_{b}\ne1$ implies that the buoyancy of the fluid
transferred depends on the properties of the receiving fluid which
is not logical and can lead to unbounded values of $b_{i}$ in the
fluid that is loosing mass (as demonstrated in section \ref{sec:results}).

\subsection{Dimensional Analysis\label{subsec:dimAnal}}

Dimensional analysis can guide our choice of $\gamma$ for the parameterisation
of the pressure difference between fluids. We make the following assumptions:
\begin{enumerate}
\item The flow is buoyancy dominated.
\item The flow is close to inviscid.
\item The flow is slowly varying.
\item Fluid $i$ properties are anomalies from a neutrally stable resting
mean state.
\item We consider only the vertical direction where $w_{ij}^{T}=0$.
\item $P_{i}=P-\gamma\nabla\cdot\mathbf{u}_{i}\approx-\gamma\nabla\cdot\mathbf{u}_{i}=-\gamma\frac{\partial w_{i}}{\partial z}$
\item There are two fluids, $i$ and $j$
\item Without loss of generality we assume that $S_{ij}=-\frac{\partial w_{i}}{\partial z}>0$
and $S_{ji}=0$.
\end{enumerate}
This leads to the following balance in the momentum equation:
\begin{equation}
\frac{\partial w_{i}^{2}}{\partial z}-\gamma\frac{\partial^{2}w_{i}}{\partial z^{2}}=b_{i}\label{eq:wi_balances}
\end{equation}
We non-dimensionalise using a length scale $L$, a buoyancy scale
$B$ and a time scale $T$ to get the non-dimensional variables:
\begin{eqnarray*}
\tilde{w} & = & w_{i}T/L\\
\tilde{b} & = & b_{i}/B\\
\tilde{z} & = & z/L.
\end{eqnarray*}
Then the non-dimensional version of (\ref{eq:wi_balances}) is:
\begin{equation}
\underbrace{{\frac{L}{T^{2}}\frac{\partial\tilde{w}^{2}}{\partial\tilde{z}}}}_{\text{advection}}-\underbrace{{\frac{\gamma}{TL}\frac{\partial^{2}\tilde{w}}{\partial\tilde{z}^{2}}}}_{\text{stabilisation}}=\underbrace{B\tilde{b}}_{\text{buoyancy}}.\label{eq:wi_nonDomTmp}
\end{equation}
The flow is buoyancy dominated so the buoyancy term should be $O(1)$
so we choose the scaling $B=L/T^{2}\implies T=(L/B)^{\frac{1}{2}}$
so the non-dimensional momentum equation becomes:
\begin{equation}
\underbrace{{\frac{\partial\tilde{w}^{2}}{\partial\tilde{z}}}}_{\text{advection}}-\underbrace{{\frac{\gamma}{B^{\frac{1}{2}}L^{\frac{3}{2}}}\frac{\partial^{2}\tilde{w}}{\partial\tilde{z}^{2}}}}_{\text{stabilisation}}=\underbrace{\tilde{b}}_{\text{buoyancy}}.\label{eq:wi_nonDom-1}
\end{equation}
The stabilisation term must be large enough to smooth out oscillations
in $\tilde{w}$ due to advection and buoyancy but not too large to
remove all variability in $\tilde{w}$ so we need:
\begin{equation}
\gamma\sim B^{\frac{1}{2}}L^{\frac{3}{2}}.\label{eq:gammaDimAnal}
\end{equation}

\section{Numerical Solution \label{sec:numerics}}

The spatial discretisation follows exactly \citet{WM19}, solving
advective form equations using monotonic, finite volume advection
and C-grid, Lorenz staggering. The time-stepping is Crank-Nicolson
with no off-centering and with deferred correction of explicitly solved
variables. It differs from \citet{WM19} because here we solve the
Boussinesq equations and we need stable solutions for equations with
a separate pressure in each fluid. We use two outer iterations per
time step to update explicitly solved variables. For the first outer
iteration, predicted values for time $n+1$ are set to those from
time level $n$ and are given the superscript $\ell$. For the second
iteration, values at $\ell$ are set to those from the end of the
first iteration. The implicit numerical method for applying the transfers
between fluids is specific for two fluids because it involves the
inversion of $2\times2$ matrices.

The prognostic variables are the $b_{i}$ in cell centres, the $\sigma_{i}$
in cell centres and the velocity flux at cell faces, $u_{i}=\mathbf{u}_{i}\cdot\mathbf{S}$,
which is the velocity at cell faces dotted with the area vector for
each face (normal to the face). The pressures, $P$ and $p_{i}$ are
diagnostic.

The first equation to be solved each outer iteration is eqn (\ref{eq:sigma})
for each $\sigma_{i}$, operator split; first applying advection,
then correcting $\sum_{i}\sigma_{i}=1$ and then applying mass transfers:
\begin{eqnarray}
\sigma_{i}^{\prime} & = & \sigma_{i}^{n}-\frac{\Delta t}{2}\nabla\cdot\left\{ \left(\mathbf{u}_{i}^{n}+\mathbf{u}_{i}^{\ell}\right)\left(\sigma_{i}^{n}\right)_{vL}\right\} \\
\sigma_{i}^{\prime\prime} & = & \sigma_{i}^{\prime}\bigg/\sum_{i}\sigma_{i}^{\prime}\\
S_{ij} & = & \begin{cases}
-\nabla\cdot\mathbf{u}_{i}^{\ell} & \text{if }\nabla\cdot\mathbf{u}_{i}^{\ell}<0\\
0 & \text{otherwise}
\end{cases}\\
\sigma_{i}^{n+1} & = & \sigma_{i}^{\prime\prime}+\Delta t\sum_{j\ne i}\sigma_{j}^{\prime\prime}S_{ji}-\sigma_{i}^{\prime\prime}S_{ij}
\end{eqnarray}
where $\Delta t$ is the time step and the superscripts $\prime$
and $\prime\prime$ denote intermediate values. $\sigma_{i}^{n}$
is interpolated from cell centres onto cell faces using monotonic
van-Leer advection (operator $(\sigma_{i}^{n})_{vL}$) as in \citet{WM19}
and values exclusively at time level $n$ for best accuracy and guaranteed
monotonicity. For calculating the divergence, the normal component
of $\mathbf{u}_{i}$ is needed at cell faces which are prognostic
variables of the C-grid. The $S_{ij}$ are limited to ensure $0<\sigma_{i}<1\ \forall i$.

Next eqn (\ref{eq:b}) is solved for the buoyancy in each fluid, first
transporting buoyancy and then applying the transfer terms. The transfer
terms can be very large due to the presence of $\frac{\sigma_{j}}{\sigma_{i}}S_{ij}$
in the transfer term for $b_{i}$ so they are treated operator split
and implicitly:
\begin{eqnarray}
b_{i}^{\prime} & = & b_{i}^{n}-\frac{\Delta t}{2}\Biggl(\left(\nabla\cdot\left(\mathbf{u}_{i}(b_{i})_{vL}\right)+b_{i}\nabla\cdot\mathbf{u}_{i}+\alpha\nabla^{2}b_{i}\right)^{n}\label{eq:transportb}\\
 &  & \hfill+\left(\nabla\cdot\left(\mathbf{u}_{i}(b_{i})_{vL}\right)+b_{i}\nabla\cdot\mathbf{u}_{i}+\alpha\nabla^{2}b_{i}\right)^{\ell}\Biggr)\\
b_{0}^{n+1} & = & b_{0}^{\prime}+\Delta t\left\{ \left(\frac{\sigma_{1}}{\sigma_{0}}\right)^{\prime\prime}S_{10}\left(b_{1}^{n+1}+b_{10}^{t}-b_{0}^{n+1}\right)-S_{01}b_{01}^{t}\right\} \label{eq:b0np1}\\
b_{1}^{n+1} & = & b_{1}^{\prime}+\Delta t\left\{ \left(\frac{\sigma_{0}}{\sigma_{1}}\right)^{\prime\prime}S_{01}\left(b_{0}^{n+1}+b_{01}^{t}-b_{1}^{n+1}\right)-S_{10}b_{10}^{t}\right\} \label{eq:b1np1}
\end{eqnarray}
where $b_{ij}^{t}=b_{ij}^{T}-b_{i}$. The use of $(\sigma_{0}/\sigma_{1})^{\prime\prime}$
from before the mass transfers is necessary to ensure bounded and
conservative transfers, as described by \cite{MWH1x}. Simultaneous
solutions of (\ref{eq:b0np1}) and (\ref{eq:b1np1}) give:
\begin{eqnarray}
b_{0}^{n+1} & = & \frac{\left(1+T_{01}\right)\left(b_{0}^{\prime}-\Delta tS_{01}b_{01}^{t}+T_{10}b_{10}^{t}\right)+T_{10}\left(b_{1}^{\prime}+T_{01}b_{01}^{t}-\Delta tS_{10}b_{10}^{t}\right)}{1+T_{01}+T_{10}}\\
b_{1}^{n+1} & = & \frac{b_{1}^{\prime}+T_{01}b_{0}^{n+1}+T_{01}b_{01}^{t}-\Delta tS_{10}b_{10}^{t}}{1+T_{01}}
\end{eqnarray}
where $T_{ij}=\Delta t\left(\frac{\sigma_{i}}{\sigma_{j}}\right)^{\prime\prime}S_{ij}$

Next a Poisson equation is solved to calculate the total pressure,
$P$ to ensure that the total velocity is divergence free. The pressure
equation also provides updates for the velocity flux, $u_{i}$. First
an intermediate velocity flux is calculated with partial updates and
updates from a previous value of $p_{i}$:
\begin{eqnarray}
u_{i}^{\prime}= & u_{i}^{n}+\frac{\Delta t}{2} & \biggl(\left(-\nabla\cdot(\mathbf{u}_{i}\mathbf{u}_{i})+\mathbf{u}_{i}\nabla\cdot\mathbf{u}_{i}+b_{i}\mathbf{k}\right)_{f}^{\ell}\cdot\mathbf{S}-\nabla_{f}p_{i}^{\ell}\\
 &  & +\left(-\nabla\cdot(\mathbf{u}_{i}\mathbf{u}_{i})+\mathbf{u}_{i}\nabla\cdot\mathbf{u}_{i}+b_{i}\mathbf{k}\right)_{f}^{n}\cdot\mathbf{S}-\nabla_{f}(P+p_{i})^{n}\biggr)
\end{eqnarray}
where operator $()_{f}$ means linear interpolation from cell centres
onto cell faces and $\nabla_{f}P$ is a compact discretisation of
$(\nabla P)_{f}\cdot\mathbf{S}$: ie the normal component of the pressure
gradient calculated from just the values of pressure either side of
the face. Once we have calculated $P^{n+1}$, the velocity flux (without
momentum transfers) is updated from the back-substitution:
\begin{equation}
u_{i}^{\prime\prime}=u_{i}^{\prime}-\frac{\Delta t}{2}\nabla_{f}P^{n+1}.\label{eq:backSub}
\end{equation}
The Poisson equation for $P^{n+1}$ is found by multiplying each eqn
(\ref{eq:backSub}) by $\sigma_{i}$, summing over all fluids and
taking the divergence. Knowing that $\nabla\cdot\sum_{i}\sigma_{i}\mathbf{u}_{i}=0$,
$\sum_{i}\sigma_{i}=1$ and $\sum_{i}\sigma_{i}p_{i}=0$ gives:
\begin{equation}
\nabla\cdot\sum_{i}\sigma_{i}u_{i}^{\prime}-\frac{\Delta t}{2}\nabla^{2}P^{n+1}=0
\end{equation}
which is solved to find $P^{n+1}$.

Next, the momentum transfers due to mass transfer and due to drag
are calculated implicitly, similar to the implicit transfers of buoyancy,
calculating $u_{i}^{n+1}$ from $u_{i}^{\prime\prime}$. These do
not influence the total $P$ or total divergence because momentum
is transferred conservatively. However they do influence $p_{i}$
which is why they are calculated before $p_{i}$. The final stage
of the time step iteration is to solve Helmholtz equations for each
$p_{i}$:
\begin{equation}
\sigma_{i}^{n+1}p_{i}^{n+1}=-\gamma\nabla\cdot\left(\sigma_{if}^{n+1}u_{i}^{n+1}-\frac{\Delta t}{2}\sigma_{if}^{n+1}\nabla_{f}\left(p_{i}^{n+1}-p_{i}^{\ell}\right)\right).\label{eq:pi_Helm}
\end{equation}
Both $p_{i}^{n+1}$ and $p_{i}^{\ell}$ appear in eqn (\ref{eq:pi_Helm})
because $u_{i}^{n+1}$ already contains contributions from $p_{i}^{\ell}$
so these are removed before updating $p_{i}^{n+1}$. No back substitution
for $u_{i}$ is done after calculating $p_{i}^{n+1}$ because this
would destroy the divergence free constraint. Therefore $p_{i}^{n+1}$
only satisfies eqn (\ref{eq:Pi_div}) exactly if the outer iterations
converge and $p_{i}^{\ell}=p_{i}^{n+1}$.

The final stage of the time step iteration is to reconstruct $\mathbf{u}_{i}$
from $u_{i}$ following \citet{WM19}. The stages calculating $P$,
$p_{i}$ and $\mathbf{u}_{i}$ are repeated twice per time step iteration.
The Poisson and Helmholtz equations are solved to a tolerance of $10^{-6}$.

\section{Validation Test Cases\label{sec:results}}

Two test cases are developed to tune parameters and to validate the
use of multi-fluid equations to represent dry, two-dimensional, sub-grid-scale
convection. The first is transient; the standard rising bubble of
\citet{BF02}. We aim to reproduce the vertical heat transport, the
mean velocity of the rising air, the bouyancy of the rising air and
pressure differences between two fluids in a single column two fluid
model. The second test case is steady state with heat transfer at
the ground and uniform radiative cooling. For both test cases the
reference solution is a single fluid solution of the two-dimensional
Boussinesq equations at high horizontal and vertical resolution.

\subsection{Rising Bubble}

>>>>

%% LyX 2.2.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage[authoryear]{natbib}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amsmath}

@@

\usepackage{babel}
\begin{document}
\setlength\tabcolsep{0.5pt}%
\begin{tabular}{ccccc}
$\sigma$ & $\sigma S\ (\text{s}^{-1})$  & $b\ (\text{m}\ \text{s}^{-2})$ & $w\ (\text{m}\ \text{s}^{-1})$ & $P\ (\text{m}^{2}\text{s}^{-2})$\tabularnewline
\multicolumn{5}{c}{Horizontal averages of the resolved RCE test case between $t=10^{5}\text{s}$
and $2\times10^{5}\text{s}$}\tabularnewline
\includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/hiRes_wide/timeMean_w/sigmaMean} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/uniformTransfer_Pi4e3_w/0/massTransfer} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/hiRes_wide/timeMean_w/bMean} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/hiRes_wide/timeMean_w/wMean} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/hiRes_wide/timeMean_w/PMean}\tabularnewline
\multicolumn{5}{c}{Single column (solid) with $S_{ij}=-\nabla\cdot\mathbf{u}_{i}$ and
$b_{ij}^{T}=b_{i}$, $\gamma=2\times10^{3}\text{m}^{2}\text{s}^{-1}$
versus resolved (dashed)}\tabularnewline
\includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_w/100000/sigma_1} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_w/100000/massTransfer} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_w/100000/b_1} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_w/100000/u_1} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_w/100000/P_1}\tabularnewline
\multicolumn{5}{c}{Single column (solid) with $S_{ij}=-\nabla\cdot\mathbf{u}_{i}$ and
$b_{ij}^{T}=\frac{1}{2}\left(b_{i}+b_{j}\right)$, $\gamma=2\times10^{3}\text{m}^{2}\text{s}^{-1}$
versus resolved (dashed)}\tabularnewline
\includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_bT05_w/100000/sigma_1} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_bT05_w/100000/massTransfer} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_bT05_w/100000/b_1} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_bT05_w/100000/u_1} & \includegraphics[scale=0.7]{/home/hilary/OpenFOAM/hilary-7/run/hilary/RCE/Boussinesq/singleColumn/divTransfer_Pi4e3_divSigmau_bT05_w/100000/P_1}\tabularnewline
\end{tabular}
\end{document}
